---
title: "A Geospatial Model of Political Ideology in California"
format: gfm
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
library(targets)
library(ggplot2)
```

In this project I estimate county-level political ideology in California using a new model I develop called **IRT-MRP-BYM**---*Item-Response-Theory, Multilevel-Regression-with-Poststratification, Besag-York-Mollié* model.[^1] 

[^1]: I'm open to any suggestions on a catchier name.

**Item Response Theory (IRT)** models are used to measure latent characteristics of individuals, such as political ideology, by using the individuals' observed choices on survey items. I use thirteen questions on abortion (2), the environment (3), gun control (3), healthcare (2), and immigration (3) from 7,998 respondents in the combined 2019, 2020, and 2021 [Cooperative Election Survey sample](https://cces.gov.harvard.edu/). 

**Multilevel Regression with Poststratification (MRP)** is a method for dealing with biased/unrepresentative survey samples. The first step involves building a multilevel model to predict the quantity of interest (ideology in this case). Then we poststratify these predictions using Census data at the county level---which has the effect of weighting the results by the prevalence of different demographic combinations in each county.

**Besag-York-Mollié (BYM)** models are used to account for geospatial dependencies in data. The basic idea is that counties which are close to one another likely share similar political ideology compared to counties which are further apart. BYM models provide a principled method for incorporating this spatial autocorrelation.[^2]

[^2]: See [Gao, Kennedy, Simpson, and Gelman 2020](https://arxiv.org/abs/1908.06716) for using BYM in MRP models.

## Results

Below are the results of the full **IRT-MRP-BYM** model for the county ideology parameter $\theta_c^{MRP}$. The coastal counties of San Mateo, Marin, and Ventura are among the most liberal, whereas the most conservative counties are in the Central Valley (Kern, Merced).

```{r}
tar_load(county_map)
county_map[[1]] +
  labs(title = "IRT-MRP-BYM model")
```

In contrast to the $\theta_c^{MRP}$ model estimates, if we were to construct a simple additive scale of ideology based on the thirteen survey questions (scaled), we get a very different picture of Californian politics. According to the additive index, Mariposa (dark green in the center of the map) is the most liberal county in the state. This defies common sense for a county which has voted for the Republican candidate every year since 1992.[^3]

[^3]: MIT Election Data and Science Lab, 2018, "County Presidential Election Returns 2000-2020", <https://doi.org/10.7910/DVN/VOQCHQ>, Harvard Dataverse, V12.

```{r}
county_map[[2]] +
  labs(title = "Additive Index of Survey Questions")
```

## Notes

- For the full paper: [paper.qmd](https://github.com/bwilden/spatial-irt/blob/main/paper.pdf)
- For the full model code in Stan: [ideal_mrp.stan](https://github.com/bwilden/spatial-irt/blob/main/stan/ideal_mrp.stan)
- For R code and cleaning steps: [_targets.R](https://github.com/bwilden/spatial-irt/blob/main/_targets.R) and [R files](https://github.com/bwilden/spatial-irt/tree/main/R)
